---
title: "Data Science I: HW3"
author: "Sydney Musick"
output: github_document
---

# Problem 1

### Load packages
```{r packages}
library(tidyverse)
library(dplyr)
```

### Load data
```{r data load}
library(p8105.datasets)
data("instacart")
```
The `instacart` dataset has `r nrow(instacart)` observations (`r nrow(instacart)` rows and `r ncol(instacart)` columns). It includes variables `r names(instacart)`. The first observation in the dataset, for example, is product_id 49302, Bulgarian Yogurt, in the yogurt aisle (aisle_id 120).

### How many aisles are there? Which is most ordered from?
```{r aisles}
max(instacart$aisle_id)
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```
There are 134 aisles. The most commonly ordered from aisles are fresh vegetables, fresh fruits, and packaged vegetables fruits.

### Plot that shows the number of items ordered in each aisle
```{r items/aisle plot}
instacart_10000 = instacart %>%
  select(aisle, department) %>%
  count(aisle, department) %>% 
  arrange(desc(n)) %>%
  filter(n >= 10000) %>%
  arrange(department) %>%
  ggplot(aes(x = aisle, y = n)) + geom_point()
  ggsave("aisle_scatter_plot.pdf", height = 4, width = 6)
```
Make a plot that shows the number of items ordered in each aisle Arrange aisles sensibly, and organize your plot so others can read it.

